Do larger watersheds respond different to forest cover change? Re-analysing a global data set.

### 2021-12-02 paper/Forest_and_Water/Forest_and_Water.Rmd
updates paper 03122021

The work by @filoso2017 and earlier by @jackson2005 has indicated that the length of the study might influence the response. This links to the idea from @kuczera1987 that the effect of logging or deforestation or reforestation reduces with the length of time post intervention (see also @jackon2005). In addition to adding the length as a variable, all continuous variables (Dryness, Area, length and Latitude) were considered non-linear in this model and as indicated a shrinkage smoothing spline [@wood2006] was applied to these variables.

@@

@@

```{r, eval =T, echo =F}
Zhang_all2 %>%
  ggplot(aes(DeltaQf_perc, length)) + 
  geom_point(colour = "blue") + theme_bw()
```

@@

```{r, echo = F, eval =F}
Zhang_all2 %>%
  ggplot(aes(DeltaF_perc_pos, DeltaQf_perc, colour = Dryness)) + geom_point()
```

```{r, echo = F}
Zhang_all2 %>%
  ggplot(aes(Hydrological_regime,Forest_type)) + geom_count() + theme_bw()
```

```{r, eval =T, echo =F}
Zhang_all2 %>%
  ggplot(aes(Assessment_technique,log10(Area_km2))) + 
  geom_boxplot(colour = "blue") +
  theme_bw() + xlab("assessment technique") + ylab("log10(Area in km2)")
```

```{r, echo = F}
Zhang_all2 %>%
  ggplot(aes(DeltaF_perc_pos, DeltaQf_perc, colour = Assessment_technique,size = log10(Area_km2), shape = Forest_Sign)) + geom_point()
```
```{r, echo = F}
png("flow_forest_byArea.png", width = 2500, height = 1800, res = 300)
Zhang_all2 %>%
  ggplot(aes(DeltaF_perc_pos, DeltaQf_perc, colour = Assessment_technique,size = log10(Area_km2), 
             shape = Forest_Sign)) + geom_point(alpha = 0.5) +
  theme_bw() + ylab("% change in flow") +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size = rel(1.5))) +
  xlab("% change in forestry") + #scale_y_log10() +
  scale_size_continuous(name = "log10(Area in km2)") +
  scale_colour_discrete(name = "Assessment Technique") +
  scale_shape_discrete(name = "Forest cover direction")
dev.off()
```

@@

As shown earlier, most of the smaller watersheds are "real observed data" using paired watershed studies, while for larger watersheds, the analysis are mostly based on modelling approximations using either elasticity analysis (EA), Hydrological modelling (HM) or a combined use of statistical methods (SH) or quasi paired watershed analsysi (QPW), thus all providing an approximation of the effect of forestry on streamflow rather than a direct comparison of watersheds. This is a confounding factor that is not easily addressed in the regression modelling attempted here. Furthermore, the catchments analysed using EA, are concentrated in the drier end of the Dryness index scale compared to the other methods, with only the paired watershed experiment (PWE) assessment technique covering the full range of dryness indices.

There are further confouding factors in the data, which were also classified by @filoso2017 and thesse create biases in the data set that can impact the overall assessment. For example, snow dominated hydrological regimes (SD), which are weakly significant, are dominated by Coniferous Forests (CF), while the majority of the rain dominated regimes are all broadleaf forests (BF). However, the forest type classification is very coarse and does not fully capture possible physiological differences that could affect evapotranspiration and therefore changes in streamflow.

>>>>

The work by @filoso2017 and earlier by @jackson2005 has indicated that the length of the study might influence the response. This links to the idea from @kuczera1987 that the effect of logging or deforestation or reforestation reduces with the length of time post intervention (see also @jackson2005). In addition to adding the length as a variable, all continuous variables (Dryness, Area, length and Latitude) were considered non-linear in this model and as indicated a shrinkage smoothing spline [@wood2006] was applied to these variables.

@@

```{r, eval =T, echo =F}
All_data2 %>%
  group_by(Assessment_technique) %>%
  summarise(n = n())
```

Maybe remove data for low numbers of assessment techniques and rerun the analysis (<10)

```{r, eval =T, echo =F}
All_data2 %>%
  ggplot(aes(Assessment_technique, log10(Area_km2))) + 
  geom_boxplot(colour = "blue") + theme_bw()
```

@@

```{r, eval =T, echo =F}
All_data2 %>%
  ggplot(aes(DeltaQf_perc, length)) + 
  geom_point(colour = "blue") + theme_bw()
```

@@

```{r, echo = F}
png("flow_forest_byArea.png", width = 2500, height = 1800, res = 300)
All_data2 %>%
  ggplot(aes(DeltaF_perc_pos, DeltaQf_perc, colour = Assessment_technique,size = log10(Area_km2), 
             shape = Forest_Sign)) + geom_point(alpha = 0.5) +
  theme_bw() + ylab("% change in flow") +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size = rel(1.5))) +
  xlab("% change in forestry") + #scale_y_log10() +
  scale_size_continuous(name = "log10(Area in km2)") +
  scale_colour_discrete(name = "Assessment Technique") +
  scale_shape_discrete(name = "Forest cover direction")
dev.off()
```

@@

As shown earlier, most of the smaller watersheds are "real observed data" using paired watershed studies, while for larger watersheds, the analysis are mostly based on modelling approximations using either elasticity analysis (EA), Hydrological modelling (HM) or a combined use of statistical methods (SH) or quasi paired watershed analysis (QPW), thus all providing an approximation of the effect of forestry on streamflow rather than a direct comparison of watersheds. This is a confounding factor that is not easily addressed in the regression modelling attempted here. Furthermore, the catchments analysed using EA, are concentrated in the drier end of the Dryness index scale compared to the other methods, with only the paired watershed experiment (PWE) assessment technique covering the full range of dryness indices.

There are further confouding factors in the data, which were also classified by @filoso2017 and these create biases in the data set that can impact the overall assessment. For example, snow dominated hydrological regimes (SD), which are weakly significant, are dominated by Coniferous Forests (CF), while the majority of the rain dominated regimes are all broadleaf forests (BF). However, the forest type classification is very coarse and does not fully capture possible physiological differences that could affect evapotranspiration and therefore changes in streamflow.
